---
title: "Midterm_Project_T6"
date: "2023-10-20"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Brew: Brewing Success with Starbucks Customer Insights

The data is contained in three files:

portfolio.csv - data about offers sent to customers (10 offers x 6 columns)

profile.csv - demographic data of customers (17,000 customers x 5 columns)

transcript.csv - customer response to offers and transactions made (306,648 events x 4 columns)

```{r, results='markup'}
# Importing and reading the Datasets 
df_cust_offer<-read.csv("data/transcript.csv")
# df_cust_offer <- df_cust_offer[, -1]
head(df_cust_offer)
print("The structure of Transcript:")
str(df_cust_offer)

df_offers<-read.csv("data/portfolio.csv")
df_offers<- df_offers[, -1]
head(df_offers)
print("The structure of Portfolio:")
str(df_offers)

df_cust<-read.csv("data/profile.csv")
df_cust<- df_cust[, -1]
head(df_cust)
print("The structure of Profile:")
str(df_cust)
```

```{r, results='markup'}
#Checking for the Missing values in each column of the dataframes
cust_offer_missing_values <- colSums(is.na(df_cust_offer))
print("Missing values in transcript:")
print(cust_offer_missing_values)


offers_missing_values <- colSums(is.na(df_offers))
print("Missing values in portfolio:")
print(offers_missing_values)


df_cust$gender[df_cust$gender == ""] <- NA #Filling empty spaces with NA values
cust_missing_values <- colSums(is.na(df_cust))
print("Missing values in profile:")
print(cust_missing_values)
```

```{r}
missing_gender <- df_cust[is.na(df_cust$gender), ]
missing_income <- df_cust[is.na(df_cust$income), ]
sum(missing_gender$id == missing_income$id)
```

```{r}
#Checking count of missing values if they are from the same observations
missing_gender_and_income <- sum(is.na(df_cust$gender) & is.na(df_cust$income))
print("Count of missing values in both gender and income:")
print(missing_gender_and_income)
```

```{r}
# Identify 'id' values to remove
id_to_remove <- missing_income$id

# Create a new dataframe without the rows where 'id' is in 'id_to_remove'
df_customer_no_na <- df_cust[!df_cust$id %in% id_to_remove, ]

# Reset the row indices
rownames(df_customer_no_na) <- NULL

# Display information about the resulting dataframe
str(df_customer_no_na)
#There are no NA values in df_customer_no_na
sum(is.na(df_customer_no_na))
```

```{r}
missing_gender_and_income_together <- df_cust[is.na(df_cust$gender) & is.na(df_cust$income), ]
head(missing_gender_and_income_together)
```

```{r}
#Checking the columns with numerical data 
cust_numcols <- names(df_customer_no_na)[sapply(df_customer_no_na, is.numeric)]
head(cust_numcols)
```

```{r}
#Converting became_member_on to datetime format
df_customer_no_na$became_member_on <- as.Date(as.character(df_customer_no_na$became_member_on), format = "%Y%m%d")
head(df_customer_no_na)
```

```{r}
#Drop the unnecessary first Column in transcript
# df_cust_offer <- df_cust_offer[, -1]
head(df_cust_offer)
```

```{r}
#check transcript data frame data types
df_cust_offer_dtypes <- sapply(df_cust_offer, class)
print(df_cust_offer_dtypes)
```

```{r}
#convert time in hours to number of days 
library(dplyr)
df_cust_offer <- df_cust_offer %>% 
  mutate(time = time / 24)
```

```{r}
#Rename columns 
colnames(df_cust_offer)[colnames(df_cust_offer) == "time"] <- "days_elapsed"
colnames(df_cust_offer)[colnames(df_cust_offer) == "person"] <- "customer_id"
tail(df_cust_offer)
```

## **Data Transformation:**

### **Transform `df_offers`**

Add an alias to each offer for easier recognition and referencing.

```{r}
#before transforming
head(df_offers)
```

```{r}
# Load required library
library(dplyr)
# Sort df_offer by offer_type and difficulty
df_offers <- df_offers %>%
  arrange(offer_type, difficulty) %>%
  mutate(index = row_number()) %>%
  select(-index)

# Add the offer_alias column
df_offers$offer_alias <- LETTERS[1:nrow(df_offers)]
head(df_offers)
```

### **Transform `df_cust_offer`**

```{r}
# Extract the 'customer_id' and 'event' columns
df_cust_offer_subset <- df_cust_offer[, c("customer_id", "event","days_elapsed")]
 
# Split the 'value' column into 'value_key' and 'value_data'
df_cust_offer_subset$value_key <- sapply(strsplit(df_cust_offer$value, "[:,]") , "[[", 1)
df_cust_offer_subset$value_data <- sapply(strsplit(df_cust_offer$value, "[:,]") , "[[", 2)
 
# Remove any leading/trailing white spaces
df_cust_offer_subset$value_key <- trimws(df_cust_offer_subset$value_key)
df_cust_offer_subset$value_data <- trimws(df_cust_offer_subset$value_data)
 
# Remove unwanted characters, including brackets and spaces, from 'value_key' and 'value_data'
df_cust_offer_subset$value_key <- gsub("[{}',\";]", "", df_cust_offer_subset$value_key)
df_cust_offer_subset$value_data <- gsub("[{}',\";]", "", df_cust_offer_subset$value_data)
# Display the resulting data frame
head(df_cust_offer_subset)
```

```{r}
df_cust_offer<- df_cust_offer_subset
head(df_cust_offer)
```

## **EDA on transcripts data:**

```{r}
library(ezids)
# Get value counts for the 'event' column
event_value_counts <- table(df_cust_offer$event)

# Print the value counts
xkabledply(event_value_counts)

```

```{r}
library(ggplot2)

# Create a data frame from the event counts
event_counts_df <- data.frame(event = names(event_value_counts), count = as.numeric(event_value_counts))

# Create a bar chart using ggplot
ggplot(event_counts_df, aes(x = event, y = count)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Bar Chart of Event Counts", x = "Event", y = "Count") +
  theme_minimal()+
  scale_y_continuous(breaks = seq(0, max(event_counts_df$count), by = 20000))



```

Observations:

1.  The dataset contains 138,953 transactions, which is more than four times the number of completed offers. This suggests that many customers made multiple purchases without the influence of an offer. This indicates a high level of customer loyalty, which is expected, given that Starbucks primarily offers fast-moving consumer goods such as coffee and food.

2.  During the campaign month, around 76,277 offers were sent to customers, and a remarkable 76% of these offers were viewed by customers. This high view rate reflects active customer engagement and suggests that the promotional slogans and design were effective in attracting customer attention.

3.  And 33,579 offers were completed, accounting for 58% of the viewed offers and 44% of all the offers sent. This relatively high offer completion rate is indicative of strong customer loyalty, likely indicating a substantial number of regular customers who regularly participate in Starbucks' promotions.

```{r}

#Print min and max of days elapsed
min_days_elapsed <- min(df_cust_offer$days_elapsed)
max_days_elapsed <- max(df_cust_offer$days_elapsed)

cat("Minimum days_elapsed:", min_days_elapsed, "\n")
cat("Maximum days_elapsed:", max_days_elapsed, "\n")

```

```{r}
library(ggplot2)
# Create a histogram
ggplot(df_cust_offer, aes(x = days_elapsed)) +
  geom_histogram(bins = 30, fill = "brown", color = "black") +
  labs(x = "Days", y = "Events Count", title = "Events Occurrence over Time") +
  theme_minimal() +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14), plot.title = element_text(size = 16))

```

The above plot reveals six distinct peaks of events during the campaign, primarily associated with the reception of offers, followed by subsequent events like offer views, completions, and purchases.

### **EDA on Customer Data:**

```{r}
head(df_customer_no_na)
# Save checkpoint after data cleaning
customers <- data.frame(df_customer_no_na)
```

### Univariate Analysis - `gender`

```{r}
# Calculate the wedge sizes and labels
wedge_size <- table(df_customer_no_na$gender)
wedge_label <- c("Male", "Female", "Other")

# Calculate percentages
total_count <- sum(wedge_size)
percentages <- 100 * wedge_size / total_count
percentages <- round(percentages, 1)

# Create a pie chart with improved aesthetics
colors <- c("lightblue", "lightgreen", "lightcoral")
par(mar = c(1, 1, 1, 1))  # Adjust margins for better appearance
pie(wedge_size, labels = paste(wedge_label, "\n", percentages, "%"), col = colors, main = "Gender")
legend("topright", legend = wedge_label, fill = colors, title = "Gender")

```

**Observations:** The customer base consists of a larger proportion of males (57.2%) compared to females (41.3%), with a minor representation (1.4%) from customers identifying with other genders.

### Univariate Analysis - `age`

```{r}
# Create a side-by-side boxplot and histogram
par(mfrow = c(1, 2), mar = c(5, 4, 4, 2))
boxplot(customers$age, xlab = "Age", main = "Boxplot", col = "lightblue")
hist(customers$age, xlab = "Age", main = "Histogram", col = "lightblue")

# Adjust axis label sizes
par(cex.lab = 1.5)

# Print descriptive statistics
summary(customers$age)

```

#### **Observations:**

-   The customer age range spans from 18 years as the youngest to 101 years as the oldest.

-   The distribution of customer ages appears to approximate a normal distribution, with a mean and standard deviation of approximately 54 and 17, respectively.

#### Dividing customers into age groups:

```{r}
# Define the age bins and labels
age_bins <- c(customers$age %>% min(), 26, 36, 46, 56, 66, 76, 86, customers$age %>% max())
age_labels <- c('18-25ys', '26-35ys', '36-45ys', '46-55ys', '56-65ys', '66-75ys', '76-85ys', '> 86ys')

# Create the age_group column
customers$age_group <- cut(customers$age, breaks = age_bins, labels = age_labels, include.lowest = TRUE)
```

```{r}
library(ggplot2)

# Calculate age group percentages
age_group_percentage <- round(prop.table(table(customers$age_group)) * 100, 1)

# Create the bar plot
ggplot(data = NULL, aes(x = names(age_group_percentage), y = age_group_percentage)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(x = "Age", y = "Percentage %", title = "Age Distribution") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12)) +
  geom_text(aes(label = paste0(age_group_percentage, "%")), position = position_stack(vjust = 0.5), size = 4) +
  scale_y_continuous(breaks = seq(0, 25, 5)) +
  coord_flip()
```

#### Observations:

-   The largest age group among customers is 56-65 years old, closely followed by the 46-55 years old group.

-   The third largest age group consists of customers aged 66-75 years.

-   Collectively, the top three age groups represent approximately 60% of the customer base.

-   This distribution is logical as individuals between 46 and 75 years of age tend to have more available time to visit cafes compared to younger age groups.

-   Additionally, they typically have better mobility than those above 75 years old, making them more likely to be cafe customers.

### Univariate Analysis - `income`

```{r}
# Set up the layout with 1 row and 2 columns for the plots
par(mfrow = c(1, 2), mar = c(5, 4, 4, 2))

# Create a boxplot for income
boxplot(customers$income, xlab = "Income", main = "Boxplot", col = "lightblue")

# Create a histogram for income
hist(customers$income, xlab = "Income", main = "Histogram", col = "lightblue")

# Adjust label sizes
par(cex.lab = 1.5)

# Print descriptive statistics for income
summary(customers$income)

```

#### Dividing customers into income groups

```{r}
# Create a new column `income_group`
customers$income_group <- cut(customers$income,
  breaks = c(min(customers$income), 45000, 60000, 75000, 90000, 105000, max(customers$income)),
  labels = c('30-45k', '45-60k', '60-75k', '75-90k', '90-105k', '> 105k')
)
```

```{r}
# Calculate income group percentages
income_group_percentage <- round(table(customers$income_group) / nrow(customers) * 100, 1)
```

```{r}
# Create a barplot
bp <- barplot(income_group_percentage, 
              main = "Income Group Distribution",
              xlab = "Income Group",
              ylab = "Percentage",
              col = "skyblue",
              names.arg = levels(customers$income_group),
              ylim = c(0, max(income_group_percentage) + 5)
)

# Add percentages within the bars
text(x = bp, y = income_group_percentage + 2, labels = paste(income_group_percentage, "%"), pos = 3)

# Add a legend
legend("topright", legend = paste(levels(customers$income_group), " (", income_group_percentage, "%)"), fill = "skyblue", bty = "n")
```

#### Observations:

-   Customer Income Range: The income of customers spans from 30k to 120k, with an average income of 65.4k.

-   Dominant Income Groups: The top two income groups are those earning 60-75k and 45-60k annually, collectively representing around 50% of the customer base.

-   Resemblance to Population: The income distribution among customers closely mirrors that of the general population. However, individuals with very low income might face challenges affording frequent visits to cafes, while those with extremely high income might opt for alternative coffee enjoyment methods.

### Univariate Analysis - `became_member_on`

```{r}
library(dplyr)
library(lubridate)

# Create a new data frame df_member
df_member <- customers %>%
  select(id, became_member_on)

# Add columns `year`, `month`, `day_of_week`
df_member <- df_member %>%
  mutate(
    year = year(became_member_on),
    month = month(became_member_on),
    day = day(became_member_on),
    month_name = month(became_member_on, label = TRUE),
    day_of_week = wday(became_member_on, label = TRUE),
    day_name = wday(became_member_on, label = TRUE, abbr = FALSE)
  )

head(df_member, 3)
```

```{r}
# Load the ggplot2 library if not already loaded
library(ggplot2)

# Create a count plot with detailed customization and values
ggplot(data = df_member, aes(x = year, fill = factor(year))) +
  geom_bar(color = "black", width = 0.7, position = "dodge") +
  geom_text(aes(label = after_stat(count)), stat = "count", vjust = -0.5, size = 4, color = "black") +
  scale_fill_brewer(palette = "Set1") +  # Use a predefined color palette
  labs(title = "Number of New Members by Year", x = "Year", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14
        )
  )

```

#### Daily member growth:

```{r}
# Load the required libraries if not already loaded
library(dplyr)
library(tidyr)

# Create a time series with the number of new members on each day
df_member_growth <- df_member %>%
  count(became_member_on) %>%
  arrange(became_member_on) %>%
  rename(`Number of New Members` = n)

```

```{r}
# Assuming df_member_growth is already defined
# Create four phases with different daily growth

# Phase 1: '2013-july' to '2015-july'
phase_1 <- df_member_growth %>%
  filter(became_member_on >= as.Date('2013-07-01') & became_member_on <= as.Date('2015-07-31'))

len_phase_1 <- nrow(phase_1)
mean_phase_1 <- mean(phase_1$`Number of New Members`)

# Phase 2: '2015-august' to '2017-july'
phase_2 <- df_member_growth %>%
  filter(became_member_on >= as.Date('2015-08-01') & became_member_on <= as.Date('2017-07-31'))

len_phase_2 <- nrow(phase_2)
mean_phase_2 <- mean(phase_2$`Number of New Members`)

# Phase 3: '2017-august' to '2018-jan'
phase_3 <- df_member_growth %>%
  filter(became_member_on >= as.Date('2017-08-01') & became_member_on <= as.Date('2018-01-31'))

len_phase_3 <- nrow(phase_3)
mean_phase_3 <- mean(phase_3$`Number of New Members`)

# Phase 4: '2018-feb' to '2018-july'
phase_4 <- df_member_growth %>%
  filter(became_member_on >= as.Date('2018-02-01') & became_member_on <= as.Date('2018-07-31'))

len_phase_4 <- nrow(phase_4)
mean_phase_4 <- mean(phase_4$`Number of New Members`)

```

```{r}
# Load the ggplot2 library if not already loaded
library(ggplot2)

# Define the means and lengths for each phase
mean_phases <- c(mean_phase_1, mean_phase_2, mean_phase_3, mean_phase_4)
len_phases <- c(len_phase_1, len_phase_2, len_phase_3, len_phase_4)

# Create a data frame with phase information
phase_info <- data.frame(
  phase = c("Phase 1", "Phase 2", "Phase 3", "Phase 4"),
  start_date = as.Date(c('2013-07-01', '2015-08-01', '2017-08-01', '2018-02-01')),
  end_date = as.Date(c('2015-07-31', '2017-07-31', '2018-01-31', '2018-07-31')),
  mean = mean_phases
)

# Create the plot with phases and mean values
ggplot(data = df_member_growth) +
  geom_line(aes(x = became_member_on, y = `Number of New Members`)) +
  geom_hline(data = phase_info, aes(yintercept = mean), linetype = "dashed", color = "red", size = 1) +
  geom_vline(data = phase_info, aes(xintercept = start_date), linetype = "dotted", color = "blue", size = 1) +
  geom_vline(data = phase_info, aes(xintercept = end_date), linetype = "dotted", color = "blue", size = 1) +
  labs(
    x = "Date",
    y = "Number of New Members",
    title = "Daily Member Growth",
    subtitle = "Phase-wise",
    caption = "Red dashed lines represent mean values for each phase"
  ) +
  scale_x_date(date_labels = "%Y-%b", date_breaks = "3 months") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20, face = "bold"),
    axis.title = element_text(size = 16),
    axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 14),
    legend.position = "none"  # Hide the legend
  )

```

#### Observations:

-   The dataset includes customers who became members between July 2013 and July 2018.

-   Only a small percentage of customers (approximately 2.2%) chose to become members from 2013 to the middle of 2015.

-   The number of new members started to increase significantly from the middle of 2015 and experienced a substantial growth spike from the middle of 2017. This suggests that the company likely executed successful campaigns during these periods, resulting in a significant boost in member growth.

-   However, in early 2018, there was a noticeable decline in daily member growth, with a 31% decrease (from 24.4 to 16.8). This drop may indicate that new strategies implemented during that time had a negative impact on member growth.

#### Member growth by Month:

```{r}
# Load the ggplot2 library if not already loaded
library(ggplot2)

# Create a detailed and rainbow-colored count plot
ggplot(data = df_member, aes(x = month_name)) +
  geom_bar(fill = rainbow(length(unique(df_member$month_name)), s = 0.7, v = 0.8), color = "black", width = 0.7) +
  labs(
    title = "Number of New Members by Month",
    x = "Count",
    y = "Month"
  ) +
  scale_fill_identity() +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14)
  )

```

#### Member growth by weekday:

```{r}
# Load the ggplot2 library if not already loaded
library(ggplot2)

# Create a custom color palette
my_colors <- c("#3498db", "#e74c3c", "#2ecc71", "#f1c40f", "#9b59b6", "#e67e22", "#1abc9c")

# Create a detailed and appealing count plot by weekday
ggplot(data = df_member, aes(x = factor(day_name, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) ) +
  geom_bar(fill = my_colors, color = "black", width = 0.7) +
  labs(
    title = "Number of New Members by Weekday",
    x = "Count",
    y = "Weekday"
  ) +
  scale_x_discrete(labels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14)
  )

```

#### Observations:

-   August, October, and December experienced the highest influx of new members, with August leading the way.

-   Conversely, February had the lowest average number of new members, indicating a drop in membership during this month.

-   Weekends, unsurprisingly, had higher member growth compared to weekdays. Among weekdays, Tuesday showed the highest member growth.

-   On average, Friday had the fewest new members, making it the day with the lowest member sign-ups.

### Bivariate Analysis - `age` vs `gender`

```{r}
# Load the required libraries if not already loaded
library(dplyr)
library(ggplot2)

# Data aggregation - mean age by gender
customers_aggregated <- customers %>%
  group_by(gender) %>%
  summarize(mean_age = mean(age))

# Print the mean age by gender
print(customers_aggregated)

# Create a KDE plot for age vs. gender
ggplot(data = customers, aes(x = age, fill = gender)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Age vs Gender",
    x = "Age",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16)
  )

```

```{r}
# Load the necessary libraries if not already loaded
library(ggplot2)

# Calculate the mean age by gender
mean_age_by_gender <- aggregate(age ~ gender, data = customers, FUN = mean)

# Print the mean age by gender
print(mean_age_by_gender)

# Create a density plot with enhanced spreading on both axes
ggplot(data = customers, aes(x = age, fill = gender)) +
  geom_density(alpha = 0.6, adjust = 2) +
  labs(
    title = "Age vs Gender",
    x = "Age",
    y = "Density"
  ) +
  ylim(0, 0.04) +  # Adjust the y-axis limits
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14)
  )

```

### Bivariate Analysis - `income` vs `gender`

```{r}
# Load the necessary libraries if not already loaded
library(ggplot2)
library(scales)

# Create a custom scaling factor for the y-axis labels
scaling_factor <- 1e5  # Adjust this factor as needed

# Create a density plot for income vs gender with increased spread on both axes
ggplot(customers, aes(x = income, fill = gender)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Income vs Gender",
    x = "Income",
    y = "Density"
  ) +
  scale_x_continuous(limits = c(0, 150000), breaks = seq(0, 150000, by = 20000)) +  # Adjust x-axis limits and breaks
  scale_y_continuous(labels = scales::number_format(scale = scaling_factor), limits = c(0, 0.000020)) +  # Adjust y-axis limits
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 14)
  )

```

#### Observations:

-   The income distribution of female customers exhibits a nearly normal distribution, suggesting that female customers across the entire income range tend to enjoy the company's products.

-   Male customers, on the other hand, have a right-skewed income distribution, indicating that a larger proportion of male customers falls within the lower half of the income spectrum among the company's customer base.

-   Female customers have a significantly higher average income compared to customers of other genders. This disparity in income may be attributed to the assumption that female customers, on average, are older than customers in other gender groups. It will be examined whether this assumption holds true in the following analysis.

-   On average, female customers have an income of \$71,000, while male customers have an average income of \$61,000. Customers of other genders have an average income of \$63,000.

### Bivariate Analysis - `income` vs `age`

```{r}
head(customers)
customers$income
```

```{r}
customers$age_group
```

```{r}
# Load required libraries
library(ggplot2)
library(dplyr)

# Create a summary dataframe
summary_data <- customers %>%
  group_by(age_group) %>%
  summarize(mean_income = mean(income))

# Create the barplot
ggplot(summary_data, aes(x = age_group, y = mean_income, fill = age_group)) +
  geom_bar(stat = "identity") +
  labs(x = "Age Group", y = "Income") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability

```

#### Observations: 

-   Customers in the two younger age groups (18-35 years old) have an average annual income of approximately \$51,000 USD.

-   The middle age groups (36-55 years old) exhibit an average annual income of about \$65,000 USD.

-   All age groups above 56 years old display very similar average annual incomes, hovering around \$70,000 USD.

-   On average, older customers of the company tend to have higher incomes.

-   The company's customers across all age groups have substantially higher average incomes compared to the median income of US citizens in 2018, which stood at \$30,119 USD.

-   The use of the 2018 median income as a benchmark is justified because the dataset contains records only up to July 2018, and customer profile data, including income, is typically not frequently updated. Hence, it's more relevant to compare customer income with the population's income in the year they last became members.
